Redshift evolution of the underlying type Ia supernova stretch distribution

### 2021-03-05 Nicolas.tex
Implemented changes and put own modifications in bf

\newcommand{\LEt}[1]{\textcolor{red}{\textbf{ #1}}}
\newcommand{\orcid}[1]{
\href{https://orcid.org/#1}{
\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}

@@

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties arises, notably that of the evolution of SN~Ia populations. We study the dependence on redshift of the SN~Ia \texttt{SALT2.4}
    light-curve stretch, which is a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to be strongly correlated with the
    SN environment, notably with stellar age tracers. We modeled the underlying
    stretch distribution as a function of redshift, using the evolution of the
    fraction of young and old SNe~Ia as predicted using the SNfactory dataset,
    and assuming a constant underlying stretch distribution for each age
    population consisting of Gaussian mixtures. We tested our prediction against
    published samples that were cut to have marginal magnitude selection effects, so that
    any observed change is indeed astrophysical and not observational in origin.
    In this first study, there are indications that the underlying SN~Ia stretch
    distribution evolves as a function of redshift, and that the young/old\LEt{the slash commonly denotes "ratio", as in "S/N" for "signal-to-noise ratio". The way you use it here is ambiguous: do you mean "and" or "or", or maybe a range?\ Please rephrase for clarity}
    drifting model is a better description of the data than any time-constant
    model, including the sample-based asymmetric distributions that are often used to
    correct Malmquist bias at a significance higher than 5 $\sigma$. The favored
    underlying stretch model is a bimodal one, composed of a high-stretch mode
    shared by both young and old environments, and a low-stretch mode that is exclusive
    to old environments. The precise effect of the redshift evolution of the intrinsic properties of a SN
    Ia population on cosmology remains to be studied. The astrophysical drift of the SN stretch distribution does affect current
    Malmquist bias corrections, however, and thereby the distances that are derived based on SN that are affected by
    observational selection effects. We highlight that this bias will increase
with surveys covering increasingly larger redshift ranges, which is particularly
important for the \LEt{please spell this out here and in the main text, you use the abbreviation only one time each}LSST.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery that the Universe expands
\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today for
understanding the properties of dark energy (DE) because it\LEt{grammatically, "it" here refers to "DE". Is this correct, or should this be "they" because you still refer to the SNe Ia, or do you mean "the probe"? If you mean the probe, please rephrase to "because this probe" for clarity} is the only tool that can
precisely map the recent expansion rate ($z<0.5$) when DE is driving it
\citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the key to directly measuring the
Hubble constant ($H_0$) if their absolute magnitude can be calibrated
\citep{riess2016, freedman2019}. Interestingly, the value of $H_0$ tha tis derived when
the SNe~Ia are anchored to Cepheids \citep[the\LEt{please provide the spelled-out versions of all instruments and surveys at first occurrence. If this interrupts the flow of the sentence too much, it may be done as a footnote} SH0ES
project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
luminosity is anchored at intermediate redshift by the baryon acoustic
oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
branch technique in place of the Cepheids seems to favor an intermediate value of
$H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
lensing also appear to favor high $H_0$ values \citep{wong2019}.

@@

The relation between SNe~Ia and their host galaxy properties has been studied
extensively. The first key finding was that the standardized SNe~Ia magnitudes
significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass
host galaxies being brighter on average \cite[e.g.,][]{kelly2010, sullivan2010,
childress2013, betoule2014, kim19, smith2020}. This mass-step correction is
currently used in cosmological analyses \citep[e.g.,][]{betoule2014,
scolnic2018a}, including to derive $H_0$ \citep{riess2016, riess2019}. The underlying connection between the SNe and their host galaxies remains
unclear, however, particularly when global properties such as the host stellar
mass are used. This in turn raises the question of the accuracy of such corrections because
the global properties of the host galaxies evolve with redshift. More recently, studies have
used the local SN host galaxy environment to probe more direct connections
between the SNe and their host galaxy environments \citep{rigault2013}, showing
that local age tracers such as the local specific star formation rate (LsSFR) or
the local color are more strongly correlated with the standardized SN magnitude
\citep{roman2018, kim18, rigault2020}. The identification of SN~Ia spectral features that are correlated with LsSFR
\citep{nordin2018} further support this connection. These results suggest age as the driving parameter
underlying the mass step. If this is true, itwould have a significant effect for
cosmology because the environmental correction that would need to be applied to SN standardization
could strongly vary with redshift \citep{rigault2013, childress2014,
scolnic2018a}. The importance of local SN environmental studies remains
highly debated, however \cite[e.g.,][]{jones2015, jones2019}, and especially the effect of
such an astrophysical bias has on the derivation of $H_0$ \citep{jones2015,
riess2016, riess2018, rose2019}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.\LEt{to properly include these references into the main text, please separate them by commas and set off the last reference additionally with "and". This is a LaTeX command error that I cannot fix for you with the program I work with (citet and citep)}
\cite{mannucci2005, scannapieco2005, sullivan2006, smith2012} have shown that
the relative SNe~Ia rate in galaxies could be explained if two populations
existed, a young population that follows the host star formation activity, and an old population that
followed the host stellar mass (the so-called prompt-and-delayed or A+B
model). \cite{rigault2020} used the LsSFR to determine the younger
(those with a high LsSFR) and the older (those with a low LsSFR) populations.
Furthermore, since the first SNe~Ia host analyses, the SN stretch has been known
to be strongly correlated with the SN host galaxy properties \citep{hamuy1996,
hamuy2000}. This correlation has been extensively confirmed since then
\citep[e.g.,][]{neill2009, lampeitl2010, gupta2011, dandrea2011, pan2014}.
Following the A+B model and the connection between SN stretch and host
galaxy properties, \cite{howell2007} first discussed the potential redshift
drift of the SN stretch distribution. In this paper we revisit this question
using the most recent SNe~Ia datasets.

@@

We present in Fig.~\ref{fig:maglim} the light-curve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008}, and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
parent distribution \citep{guy2007, betoule2014}; see also
\citet{bazin2011} and \citet{campbell2013} for similar contours, the second
using a more conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude
with $x_1=0$ and $c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can
derive the absolute standardized magnitude at maximum of light $M = M_0 - \alpha
x_1 + \beta c$ within the aforementioned ellipse given the standardization
coefficient $\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: The
faintest SN~Ia is that with $(x_1=-1.65, c=0.25)$ and an absolute standardized
magnitude at peak in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. This object typically ought to be detected 5 days before and a week after peak to
build a suitable light curve, the effective limiting standardized absolute
magnitude is approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude
limit $m_{\lim}$ of a magnitude limited survey, we can derive the maximum
redshift $z_{\lim}$ above which the faintest SNe~Ia will be missed using the
relation between apparent magnitude, redshift, and absolute magnitude
$\mu(z_{\lim}) = m_{\lim} - M_{\lim}$.

@@

\cite{rigault2020} presented a model for the evolution of the fraction of
younger and older SNe~Ia as a function of redshift following previous work on
rates and delay-time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, smith2012, childress2014, maozmannucci2014}.
In short, it was assumed that the number of young SNe~Ia follows the star
formation rate (SFR) in the Universe, while the number of old SNe~Ia follows
the number of billion-old stars in the Universe, that is, the stellar mass (M$^*$).
Hence, if we denote $\delta(z)$ ( $\psi(z) = 1-\delta(z)$) the fraction of
young (old) SNe~Ia in the Universe as a function of redshift, then the
ratio $\delta/\psi$ is expected to follow the evolution of the specific star
formation rate (SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
\citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
\citep{rigault2013, rigault2020, wiseman2020}, in agreement with rate
expectations \citep{mannucci2006, rodney2014}, \cite{rigault2020} concluded that

\begin{equation}
    \label{eq:delta}
    \delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
\end{equation}
with $K=0.87$. This model is comparable to the evolution subsequently
predicted by \cite{childress2014} based on SN rates in galaxies depending on
their quenching time as a function of their stellar mass.

@@

To model the evolution of the full SN stretch distribution as a function of
redshift, we need to model the SN stretch
distribution for each age subsample given our aforementioned model of the evolution of the fraction of
younger and older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between SN stretch and LsSFR
measurement, a progenitor age tracer, using the SNfactory sample. This relation
is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory SNe used in the
current analysis. Given the structure of the stretch-LsSFR scatter plot, our
model of the underlying SN~Ia stretch distribution is defined as follows\LEt{please reformat so that this becomes proper text, there is no need for a two-item bulleted list}:

@@

\begin{itemize}
    \item for the younger population (i.e., $\log(\mathrm{LsSFR})\geq-10.82$),
        the stretch distribution is modeled as a single normal distribution
        $\mathcal{N}(\mu_1, \sigma_1{}^2)$; 
    \item the older population (i.e., $\log( \mathrm{LsSFR})<-10.82$) stretch
        distribution is modeled as a bimodal Gaussian mixture $a\times
        \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times \mathcal{N}(\mu_2,
        \sigma_2{}^2)$, where one mode is the same as for the young population,
        $a$ representing the relative effect of the two modes.
\end{itemize}

@@

Given the probability $y^i$ that a given SN is young and assuming our base
model (see Section~\ref{sec:basemodel}), the probability of measuring a
\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg].
\end{align}

The maximum-likelihood estimate of the five~free parameters
$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
by minimizing the following:
\begin{equation}\label{eq:likelihood}
    -2\ln(L) = -2 \sum_i \ln \prob{x_1^i}{\vec{\theta};
    \mathrm{d}x_1^i, y^i}.
\end{equation}

@@

\begin{table*}
    \centering
    \caption{Best-fit values of the parameters for the base stretch distribution
    model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
569 SN~Ia sample, or the conservative model (422).}
    \label{tab:modelresults}
    \begin{tabular}{lccccc}
        \hline\hline
        Sample & $\mu_1$ & $\sigma_1$
               & $\mu_2$ & $\sigma_2$
               & $a$ \\
        \hline
        SNfactory & $ 0.41 \pm 0.08$ & $0.55 \pm 0.06$
                  & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$
                  & $ 0.48 \pm 0.08$ \\
        Fiducial & $ 0.37 \pm 0.05$ & $0.61 \pm 0.04$
                 & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$
                 & $ 0.51 \pm 0.09$ \\
        Conservative & $ 0.38 \pm 0.05$ & $0.60 \pm 0.04$
                     & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$
                     & $ 0.47 \pm 0.09$ \\
        \hline
    \end{tabular}
\end{table*}

@@

When direct LsSFR measurements are lacking (i.e., $p_y^i$), we can extend the
analysis to non-SNfactory samples by using the redshift evolution of the
fraction $\delta(z)$ of young SNe~Ia (Eq.~\ref{eq:delta}) as a proxy for the
probability of a SN to be young. This still corresponds to minimizing
Eq.~\ref{eq:likelihood} with respect to the parameters
$\vec{\theta}\equiv(\mu_1, \mu_2, \sigma_1, \sigma_2, a)$ of the stretch
distribution $X_1$ (Eq.~\ref{eq:stretchz}), but this time, assuming $y^i =
\delta(z^i)$ for any given SN~$i$. 
\\
\\
For the remaining analysis, we therefore minimized Eq.~\ref{eq:likelihood}
using $p_y^i$ , the probability for the SN $i$ to be young, when available
(i.e., for SNfactory dataset), and $\delta(z^i),$ the expected fraction of
young SNe~Ia at the SN redshift $z^i$  , otherwise.

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(conservative) sample are given Table~\ref{tab:modelresults}, and the
predicted redshift evolution of mean stretch (expected $x_1$ given the
distribution of Eq.~\ref{eq:stretchz}) illustrated as a blue band in
Fig.~\ref{fig:modelall} accounting for parameter errors and their covariances.
This figure shows that the measured mean SN~Ia stretch per redshift bin of
equal sample size closely follows our redshift drift modeling. This is indeed
what is expected if old environments favor low SN stretches
\citep[e.g.,][]{howell2007} \textit{\textup{and}} if the fraction of old SNe~Ia declines as
a function of redshift. See Section~\ref{sec:results} for a more quantitative
discussion.

@@

Because the various models have different degrees of freedom, we used the Akaike
information criterion \citep[AIC, e.g.,][]{burnham2004} to compare their ability
to properly describe the observations. This estimator penalizes additional degrees of
freedom to avoid overfitting the data and is defined as follows:
\begin{equation}
    \mathrm{AIC} = -2\ln(L) + 2k
,\end{equation}
where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
the number of free parameters to be adjusted. The reference model has the smallest AIC; in comparison to this model, the models with
$\Delta\mathrm{AIC}<5$ are coined acceptable, those with
$5<\Delta\mathrm{AIC}<20$ are not favored, and those with $\Delta\mathrm{AIC}>20$
are deemed excluded. This roughly corresponds to 2, 3, and 5~$\sigma$ limits for
a Gaussian probability distribution.

@@

We note that when models are compared that were adjusted on individual subsamples rather
than globally, the Bayesian information criterion ($\mathrm{BIC} = -2\ln(L) +
k\ln(n)$, with $n$ the number of data points) might be better suited than AIC because it explicitly accounts for the fact that each subsample is fitted
separately: the sample-based model BIC is rightfully the sum of the BIC for each
sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the sample-based
asymmetric Gaussian model as being as pertinent as the base model.

@@

We report in Table~\ref{tab:bbc} our determination of $\mu_0$ and
$\sigma_{\pm}$ for each sample when an asymmetric Gaussian
model was implmented, and adjusted on the nominally selection-free samples using our fiducial cuts
(see Section~\ref{sec:sample}). Our results are in close agreement with those of
\cite{scolnic2016} for the SNLS and SDSS and with the results reported by \cite{scolnic2018a} for PS1, who
derived these model parameters using the full BBC formalism, using numerous
simulations to model the observational selection effects (see details,
e.g., Section~3 of \citealt{kessler2017}). The agreement between our fit of the
asymmetric Gaussians on the supposedly selection-free part of the samples and
the results derived using the BBC formalism supports our approach to
constructing a sample with negligible observational selection
effects. If we were to use the \cite{scolnic2016} and \cite{scolnic2018a} best-fit
values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for the SNLS, SDSS and
PS1, respectively, the $\Delta$AIC between our base drifting model and the BBC
modeling would go even deeper from $-32$ to $-47$. We further discuss the
consequence of this result for cosmology in Section~\ref{sec:discussion}.

@@

Finally, ignoring the LsSFR measurements, which are only available for the SNfactory
dataset (see Section~\ref{sec:modeling}), reduces the significance of the
results presented in this section, as expected. Even so, nondrifting
models remain strongly disfavored. For instance, the best-fitting sample-based Gaussian asymmetric model is still
$\Delta\mathrm{AIC}<-10,$ which is less representative of the data than our base drifting
model.

>>>>

\newcommand{\orcid}[1]{
\href{https://orcid.org/#1}{
\includegraphics[height=11pt]{General_figures/ORCIDiD_icon128x128.png}}}

@@

\abstract{The detailed nature of type Ia supernovae (SNe~Ia) remains uncertain,
    and as survey statistics increase, the question of astrophysical systematic
    uncertainties arises, notably that of the evolution of SN~Ia populations. We
    study the dependence on redshift of the SN~Ia \texttt{SALT2.4} light-curve
    stretch, which is a purely intrinsic SN property, to probe its potential
    redshift drift. The SN stretch has been shown to be strongly correlated with
    the SN environment, notably with stellar age tracers. We modeled the
    underlying stretch distribution as a function of redshift, using the
    evolution of the fraction of young and old SNe~Ia as predicted using the
    SNfactory dataset, and assuming a constant underlying stretch distribution
    for each age population consisting of Gaussian mixtures. We tested our
    prediction against published samples that were cut to have marginal
    magnitude selection effects, so that any observed change is indeed
    astrophysical and not observational in origin. In this first study, there
    are indications that the underlying SN~Ia stretch distribution evolves as a
    function of redshift, and that the \textbf{age} drifting model is a better
    description of the data than any time-constant model, including the
    sample-based asymmetric distributions that are often used to correct
    Malmquist bias at a significance higher than 5 $\sigma$. The favored
    underlying stretch model is a bimodal one, composed of a high-stretch mode
    shared by both young and old environments, and a low-stretch mode that is
    exclusive to old environments. The precise effect of the redshift evolution
    of the intrinsic properties of a SN Ia population on cosmology remains to be
    studied. The astrophysical drift of the SN stretch distribution does affect
    current Malmquist bias corrections, however, and thereby the distances that
    are derived based on SN that are affected by observational selection
effects. We highlight that this bias will increase with surveys covering
increasingly larger redshift ranges, which is particularly important for the
Large Synoptic Survey Telescope.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
enabled the discovery that the Universe\textbf{'s expansion accelerates}
\citep{riess1998, perlmutter1999}. They remain a key cosmological probe today
for understanding the properties of dark energy (DE) because they are the only
tool that can precisely map the recent expansion rate ($z<0.5$) when DE is
driving it \citep[e.g.,][]{scolnicastro2020}. Type Ia supernovae are also the
key to directly measuring the Hubble constant ($H_0$) if their absolute
magnitude can be calibrated \citep{riess2016, freedman2019}. Interestingly, the
value of $H_0$ that is derived when the SNe~Ia are anchored to Cepheids
\citep[the Supernovae, $H_0$, for the Equation of State of dark energy
project,][]{riess2009, riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\Lambda$CDM \citep{planck2018, riess2019, reid2019}, or when the SN
luminosity is anchored at intermediate redshift by the baryon acoustic
oscillation (BAO) scale \citep{feeney2019}. While using the tip of the red giant
branch technique in place of the Cepheids seems to favor an intermediate value
of $H_0$ \citep{freedman2019, freedman2020}, time-delay measurements from strong
lensing also appear to favor high $H_0$ values \citep{wong2019}.

@@

The relation between SNe~Ia and their host galaxy properties has been studied
extensively. The first key finding was that the standardized SNe~Ia magnitudes
significantly depend on the host galaxy stellar mass, with SNe~Ia from high-mass
host galaxies being brighter on average \cite[e.g.,][]{kelly2010, sullivan2010,
childress2013, betoule2014, kim19, smith2020}. This mass-step correction is
currently used in cosmological analyses \citep[e.g.,][]{betoule2014,
scolnic2018a}, including to derive $H_0$ \citep{riess2016, riess2019}. The
underlying connection between the SNe and their host galaxies remains unclear,
however, particularly when global properties such as the host stellar mass are
used. This in turn raises the question of the accuracy of such corrections
because the global properties of the host galaxies evolve with redshift. More
recently, studies have used the local SN host galaxy environment to probe more
direct connections between the SNe and their host galaxy environments
\citep{rigault2013}, showing that local age tracers such as the local specific
star formation rate (LsSFR) or the local color are more strongly correlated with
the standardized SN magnitude \citep{roman2018, kim18, rigault2020}. The
identification of SN~Ia spectral features that are correlated with LsSFR
\citep{nordin2018} further support this connection. These results suggest age as
the driving parameter underlying the mass step. If this is true, it would have a
significant effect for cosmology because the environmental correction that would
need to be applied to SN standardization could strongly vary with redshift
\citep{rigault2013, childress2014, scolnic2018a}. The importance of local SN
environmental studies remains highly debated, however \cite[e.g.,][]{jones2015,
jones2019}, and especially the effect of such an astrophysical bias has on the
derivation of $H_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

The concept of the SN~Ia age dichotomy arose with the study of the SN~Ia rate.
\cite{mannucci2005}, \cite{scannapieco2005}, \cite{sullivan2006} and
\cite{smith2012} have shown that the relative SNe~Ia rate in galaxies could be
explained if two populations existed, a young population that follows the host
star formation activity, and an old population that followed the host stellar
mass (the so-called prompt-and-delayed or A+B model). \cite{rigault2020} used
the LsSFR to determine the younger (those with a high LsSFR) and the older
(those with a low LsSFR) populations. Furthermore, since the first SNe~Ia host
analyses, the SN stretch has been known to be strongly correlated with the SN
host galaxy properties \citep{hamuy1996, hamuy2000}. This correlation has been
extensively confirmed since then \citep[e.g.,][]{neill2009, lampeitl2010,
gupta2011, dandrea2011, pan2014}. Following the A+B model and the connection
between SN stretch and host galaxy properties, \cite{howell2007} first discussed
the potential redshift drift of the SN stretch distribution. In this paper we
revisit this question using the most recent SNe~Ia datasets.

@@

We present in Fig.~\ref{fig:maglim} the light-curve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008}, and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full parent
distribution \citep{guy2007, betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is
$M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
within the aforementioned ellipse given the standardization coefficient
$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: The faintest SN~Ia is
that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
in Bessel $B$ band of $M^{t_0}_{\min} = -18.31$~mag. This object typically ought
to be detected 5 days before and a week after peak to build a suitable
light-curve, the effective limiting standardized absolute magnitude is
approximately $M_{\lim} = -18.00$~mag. Hence, given the magnitude limit
$m_{\lim}$ of a magnitude limited survey, we can derive the maximum redshift
$z_{\lim}$ above which the faintest SNe~Ia will be missed using the relation
between apparent magnitude, redshift, and absolute magnitude $\mu(z_{\lim}) =
m_{\lim} - M_{\lim}$.

@@

\cite{rigault2020} presented a model for the evolution of the fraction of
younger and older SNe~Ia as a function of redshift following previous work on
rates and delay-time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, smith2012, childress2014, maozmannucci2014}. In
short, it was assumed that the number of young SNe~Ia follows the star formation
rate (SFR) in the Universe, while the number of old SNe~Ia follows the number of
billion-old stars in the Universe, that is, the stellar mass (M$^*$). Hence, if
we denote $\delta(z)$ ($\psi(z) = 1-\delta(z)$) the fraction of young (old)
SNe~Ia in the Universe as a function of redshift, then the ratio $\delta/\psi$
is expected to follow the evolution of the specific star formation rate
(SFR/M$^*$), which goes as $(1+z)^{2.8}$ until $z\sim2$
\citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
\citep{rigault2013, rigault2020, wiseman2020}, in agreement with rate
expectations \citep{mannucci2006, rodney2014}, \cite{rigault2020} concluded that
\begin{equation}
    \label{eq:delta}
    \delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
\end{equation}
with $K=0.87$. This model is comparable to the evolution subsequently
predicted by \cite{childress2014} based on SN rates in galaxies depending on
their quenching time as a function of their stellar mass.

@@

To model the evolution of the full SN stretch distribution as a function of
redshift, we need to model the SN stretch distribution for each age subsample
given our aforementioned model of the evolution of the fraction of younger and
older SNe~Ia with cosmic time. \cite{rigault2020} presented the relation between
SN stretch and LsSFR measurement, a progenitor age tracer, using the SNfactory
sample. This relation is shown in Fig.~\ref{fig:stretchlssfr} for the SNfactory
SNe used in the current analysis. Given the structure of the stretch-LsSFR
scatter plot, our model of the underlying SN~Ia stretch distribution is defined
as follows: \textbf{the stretch distribution of the younger population
($\log(\mathrm{LsSFR})\geq-10.82$) is modeled as a single normal distribution
$\mathcal{N}(\mu_1, \sigma_1{}^2)$, and the stretch distribution of the older
population ($\log(\mathrm{LsSFR})<-10.82$) is modeled as a bimodal Gaussian
mixture $a\times \mathcal{N}(\mu_1, \sigma_1{}^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2{}^2)$, where one mode is the same as for the young
population, $a$ representing the relative effect of the two modes.}

@@

@@

Given the probability $y^i$ that a given SN is young and assuming our base model
(see Section~\ref{sec:basemodel}), the probability of measuring a
\textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error $\mathrm{d}x_1^i$ is
given by
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg].
\end{align}

The maximum-likelihood estimate of the five~free parameters
$\vec{\theta}\equiv({\mu_1, \mu_2, \sigma_1, \sigma_2,a})$ of the model is
obtained by minimizing the following:
\begin{equation}\label{eq:likelihood}
    -2\ln(L) = -2 \sum_i \ln \prob{x_1^i}{\vec{\theta};
    \mathrm{d}x_1^i, y^i}.
\end{equation}

@@

\begin{table*}
    \centering
    \caption{Best-fit values of the parameters for the base stretch distribution
    model when applied to the SNfactory dataset only (114 SNe~Ia), the fiducial
569 SN~Ia sample, or the conservative \textbf{sample} (422).}
    \label{tab:modelresults}
    \begin{tabular}{lccccc}
        \hline\hline
        Sample & $\mu_1$ & $\sigma_1$
               & $\mu_2$ & $\sigma_2$
               & $a$ \\
        \hline
        SNfactory & $ 0.41 \pm 0.08$ & $0.55 \pm 0.06$
                  & $-1.38 \pm 0.10$ & $0.44 \pm 0.08$
                  & $ 0.48 \pm 0.08$ \\
        Fiducial & $ 0.37 \pm 0.05$ & $0.61 \pm 0.04$
                 & $-1.22 \pm 0.16$ & $0.56 \pm 0.10$
                 & $ 0.51 \pm 0.09$ \\
        Conservative & $ 0.38 \pm 0.05$ & $0.60 \pm 0.04$
                     & $-1.26 \pm 0.13$ & $0.53 \pm 0.08$
                     & $ 0.47 \pm 0.09$ \\
        \hline
    \end{tabular}
\end{table*}

@@

When direct LsSFR measurements are lacking (i.e., $p_y^i$), we can extend the
analysis to non-SNfactory samples by using the redshift evolution of the
fraction $\delta(z)$ of young SNe~Ia (Eq.~\ref{eq:delta}) as a proxy for the
probability of a SN to be young. This still corresponds to minimizing
Eq.~\ref{eq:likelihood} with respect to the parameters
$\vec{\theta}\equiv(\mu_1, \mu_2, \sigma_1, \sigma_2, a)$ of the stretch
distribution $X_1$ (Eq.~\ref{eq:stretchz}), but this time, assuming $y^i =
\delta(z^i)$ for any given SN~$i$.\bigbreak

\noindent For the remaining analysis, we therefore minimized
Eq.~\ref{eq:likelihood} using $p_y^i$, the probability for the SN $i$ to be
young, when available (i.e., for SNfactory dataset), and $\delta(z^i)$, the
expected fraction of young SNe~Ia at the SN redshift $z^i$, otherwise.

Results of fitting this model to all the 569 (resp. 422) SNe from the fiducial
(conservative) sample are given Table~\ref{tab:modelresults}, and the predicted
redshift evolution of mean stretch (expected $x_1$ given the distribution of
Eq.~\ref{eq:stretchz}) illustrated as a blue band in Fig.~\ref{fig:modelall}
accounting for parameter errors and their covariances. This figure shows that
the measured mean SN~Ia stretch per redshift bin of equal sample size closely
follows our redshift drift modeling. This is indeed what is expected if old
environments favor low SN stretches \citep[e.g.,][]{howell2007} and if the
fraction of old SNe~Ia declines as a function of redshift. See
Section~\ref{sec:results} for a more quantitative discussion.

@@

Because the various models have different degrees of freedom, we used the Akaike
information criterion \citep[AIC, e.g.,][]{burnham2004} to compare their ability
to properly describe the observations. This estimator penalizes additional
degrees of freedom to avoid overfitting the data and is defined as follows:
\begin{equation}
    \mathrm{AIC} = -2\ln(L) + 2k,
\end{equation}
where $-2\ln(L)$ is derived by minimizing Eq.~\eqref{eq:likelihood}, and $k$ is
the number of free parameters to be adjusted. The reference model has the
smallest AIC; in comparison to this model, the models with
$\Delta\mathrm{AIC}<5$ are coined acceptable, those with
$5<\Delta\mathrm{AIC}<20$ are not favored, and those with
$\Delta\mathrm{AIC}>20$ are deemed excluded. This roughly corresponds to 2, 3,
and 5~$\sigma$ limits for a Gaussian probability distribution.

@@

We note that when models that were adjusted on individual subsamples rather than
globally \textbf{are compared}, the Bayesian information criterion
($\mathrm{BIC} = -2\ln(L) + k\ln(n)$, with $n$ the number of data points) might
be better suited than AIC because it explicitly accounts for the fact that each
subsample is fitted separately: the sample-based model BIC is rightfully the sum
of the BIC for each sample. We find $\Delta\mathrm{BIC}=-48$, again refuting the
sample-based asymmetric Gaussian model as being as pertinent as the base model.

@@

We report in Table~\ref{tab:bbc} our determination of $\mu_0$ and $\sigma_{\pm}$
for each sample when an asymmetric Gaussian model was implemented, and adjusted
on the nominally selection-free samples using our fiducial cuts (see
Section~\ref{sec:sample}). Our results are in close agreement with those of
\cite{scolnic2016} for the SNLS and SDSS and with the results reported by
\cite{scolnic2018a} for PS1, who derived these model parameters using the full
BBC formalism, using numerous simulations to model the observational selection
effects (see details, e.g., Section~3 of \citealt{kessler2017}). The agreement
between our fit of the asymmetric Gaussians on the supposedly selection-free
part of the samples and the results derived using the BBC formalism supports our
approach to constructing a sample with negligible observational selection
effects. If we were to use the \cite{scolnic2016} and \cite{scolnic2018a}
best-fit values of the $\mu_0, \sigma_{\pm}$ asymmetric parameters for the SNLS,
SDSS and PS1, respectively, the $\Delta$AIC between our base drifting model and
the BBC modeling would go even deeper from $-32$ to $-47$. We further discuss
the consequence of this result for cosmology in Section~\ref{sec:discussion}.

@@

Finally, ignoring the LsSFR measurements, which are only available for the
SNfactory dataset (see Section~\ref{sec:modeling}), reduces the significance of
the results presented in this section, as expected. Even so, nondrifting models
remain strongly disfavored. For instance, the best-fitting sample-based Gaussian
asymmetric model is still $\Delta\mathrm{AIC}<-10$, which is less representative
of the data than our base drifting model.
